Spectral Experts for Estimating Mixtures of Linear Regressions

### 2013-06-13 doc/icml-presentation/presentation.tex
Added crisp figures for experiments. Good version of slideS

\begin{frame}
  \frametitle{Example: Gaussian Mixture Model}
  \splitcolumn{%
    \begin{itemize}
      \item Generative process\tikzmark{gen}:
      \begin{align*}
        h &\sim \Mult([\pi_1, \pi_2, \cdots, \pi_k]) \\
        x &\sim \normal{\beta_h}{\sigma^2}.
      \end{align*}
      \item Moments:
      \begin{align*}
          \uncover<2->{%
          \E[x|h] &= \beta_h \\
          }
          \uncover<3->{%
          \E[x] &= \sum_h \pi_h \beta_h \\
          }
          \uncover<4->{%
          \tikzmark{m2}
          \E[x\tp{2}] &= \sum_h \pi_h (\beta_h \beta_h^T) \\
                      &= \sum_h \pi_h \beta_h{\tp{2}} \\
                      }
          \uncover<5->{%
          \tikzmark{m3}
          \E[x\tp{3}] &= \sum_h \pi_h \beta_h\tp{3}.
          }
        \end{align*}
    \end{itemize}
  }{%
  }
  \begin{canvas}
    % The model
    \point{mark}{(1cm,0)};
    \point{start}{(1cm,1cm)}; %{pic cs:gen} -| mark)};
    \drawgen{($(start) + (0,1cm)$)};
    %\node[anchor=west] (diag1) at ($(start)$) {%
    %  \includegraphics[width=0.45\textwidth,height=2cm,keepaspectratio]{figures/gen.png}
    %};
    \node[anchor=west] (diag) at ($(start) + (1cm,0.0cm)$) {%
      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.pdf}
    };

@@

\begin{frame}
  \frametitle{Recovering Parameters.}
  \centering

@@

\begin{frame}
  \frametitle{Spectral Experts}
  \cornertext{%
  \uncover<2->{\cite{NegahbanWainwright2009,Tomioka2011}} \\
  \uncover<3->{\cite{AnandkumarGeHsu2012}} 
  }

@@

  \begin{canvas}
    % x, y
    \node[anchor=west] (em) at (-6.5cm,-1.5cm) {%
      \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3.png}
      \includegraphics<2>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-em.png}
      \includegraphics<3>[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-spec.png}
      \includegraphics<4->[width=\textwidth,height=5cm,keepaspectratio]{figures/1-8-3-specm.png}
      };
    % x, y
    \node[anchor=west, right=6cm of em] at (em.west) {%
      \includegraphics[width=\textwidth,height=5cm,keepaspectratio]{figures/hist.png}
    };

    \node[above=0.1 cm of em] {%
        $y = \beta^T 
            \begin{bmatrix} 
              1 \\
              t \\
              t^4 \\
              t^7
            \end{bmatrix} + \epsilon$
      };
  \end{canvas}
\end{frame}

@@

\begin{small}
  \begin{tabular}{r r r c c c}
\hline
%$b$ & 
$d$ & $k$ & Spectral & EM & Spectral + EM \\
\hline
  %1 & 
  4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
  %2 & 
  5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
  %2 & 
  5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
  %2 & 
  6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
\hline
\end{tabular}
      \end{small}

@@

    \draw<5->[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
    %\node[scale=0.7] at (mle) {x};
    \node<5->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\theta^*}_{\textrm{spec + EM}}$};

  \end{tikzpicture}

@@

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
    \item<1-> Expose tensor factorization structure through regression.
    \item<2-> {\bf Consistent estimator} for the mixture of linear regressions with {\bf polynomial sample and computational complexity}.
    \item<3-> Empirically, method of moment estimates can be a good initialization for EM.
    \item<4-> How can we handle other discriminative models?
      \begin{itemize}
          \item<5-> Dependencies between $h$ and $x$ (mixture of experts).
          \item<6-> Non-linear link functions (logistic regression).
      \end{itemize}
  \end{itemize}
\end{frame}

>>>>

\begin{frame}
  \frametitle{Example: Gaussian Mixture Model}
  \splitcolumn{%
    \begin{itemize}
      \item Generative process\tikzmark{gen}:
      \begin{align*}
        h &\sim \Mult([\pi_1, \pi_2, \cdots, \pi_k]) \\
        x &\sim \normal{\beta_h}{\sigma^2}.
      \end{align*}
      \item Moments:
      \begin{align*}
          \uncover<2->{%
          \E[x|h] &= \beta_h \\
          }
          \uncover<3->{%
          \E[x] &= \sum_h \pi_h \beta_h \\
          }
          \uncover<4->{%
          \tikzmark{m2}
          \E[x\tp{2}] &= \sum_h \pi_h (\beta_h \beta_h^T) + \sigma^2 \\
                      &= \sum_h \pi_h \beta_h{\tp{2}} + \sigma^2 \\
                      }
          \uncover<5->{%
          \tikzmark{m3}
          \E[x\tp{3}] &= \sum_h \pi_h \beta_h\tp{3} + \textrm{bias}.
          }
        \end{align*}
    \end{itemize}
  }{%
  }
  \begin{canvas}
    % The model
    \point{mark}{(1cm,0)};
    \point{start}{(1cm,1cm)}; %{pic cs:gen} -| mark)};
    \drawgen{($(start) + (0,1cm)$)};
    %\node[anchor=west] (diag1) at ($(start)$) {%
    %  \includegraphics[width=0.45\textwidth,height=2cm,keepaspectratio]{figures/gen.png}
    %};
    \node[anchor=west] (diag) at ($(start) + (1cm,0.0cm)$) {%
      \includegraphics[width=0.45\textwidth,height=3cm,keepaspectratio]{figures/mog.pdf}
    };

@@

\begin{frame}
  \frametitle{Recovering Parameters}
  \centering

@@

\begin{frame}
  \frametitle{Overview: Spectral Experts}
  \cornertext{%
  \uncover<2->{\cite{NegahbanWainwright2009,Tomioka2011}} \\
  \uncover<3->{\cite{AnandkumarGeHsu2012}} 
  }

@@

  \begin{canvas}
    % x, y
    \node[anchor=west] (em) at (-6.5cm,-0.0cm) {%
    \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1833.pdf}
      \includegraphics<2-3>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-1833.pdf}
      \includegraphics<4-5>[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-1833.pdf}
      \includegraphics<6->[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-Spectral+EM-1833.pdf}
      };

    \node<1-2>[anchor=west, right=1 cm of em.east] {%
        $y = \beta^T 
            \begin{bmatrix} 
              1 \\
              t \\
              t^4 \\
              t^7
            \end{bmatrix} + \epsilon$
      };
    % x, y
    \node[anchor=west, right=6.2cm of em] at (em.west) {%
      \includegraphics<3-4>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-hist.pdf}
      \includegraphics<5>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-hist.pdf}
      \includegraphics<6>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-Spectral-Spectral+EM-hist.pdf}
    };
  \end{canvas}
\end{frame}

@@

  \begin{canvas}
    \node (exp1) at (-3cm,2cm) {\includegraphics[width=4.5cm]{figures/err-hist-0.pdf}};
    \node[below=-0.2cm of exp1] {$d=4, k=2$};
    \node (exp2) at (3cm,2cm) {\includegraphics[width=4.5cm]{figures/err-hist-1.pdf}};
    \node[below=-0.2cm of exp2] {$d=5, k=2$};
    \node (exp3) at (-3cm,-2cm) {\includegraphics[width=4.5cm]{figures/err-hist-2.pdf}};
    \node[below=-0.2cm of exp3] {$d=5, k=3$};
    \node (exp4) at (3cm,-2cm) {\includegraphics[width=4.5cm]{figures/err-hist-3.pdf}};
    \node[below=-0.2cm of exp4] {$d=6, k=2$};
  \end{canvas}

%\begin{small}
%  \begin{tabular}{r r r c c c}
%\hline
%%$b$ & 
%$d$ & $k$ & Spectral & EM & Spectral + EM \\
%\hline
%  %1 & 
%  4 & 2 & 2.45 $\pm$ 3.68 & 0.28 $\pm$ 0.82 & {\bf 0.17 $\pm$ 0.57} \\
%  %2 & 
%  5 & 2 & 1.38 $\pm$ 0.84 & {\bf 0.00 $\pm$ 0.00} & {\bf 0.00 $\pm$ 0.00} \\
%  %2 & 
%  5 & 3 & 2.92 $\pm$ 1.71 & 0.43 $\pm$ 1.07 & {\bf 0.31 $\pm$ 1.02} \\
%  %2 & 
%  6 & 2 & 2.33 $\pm$ 0.67 & 0.63 $\pm$ 1.29 & {\bf 0.01 $\pm$ 0.01} \\
%\hline
%\end{tabular}
%      \end{small}

@@

    \draw<5->[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
    \node<5->[scale=0.7] at (mle) {x};
    \node<5->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\theta^*}_{\textrm{spec + EM}}$};
    \draw<5->[dashed,blue,line width=0.7pt] ($(mle)-(0.4cm,0)$) -- ($(mle)+(0.4cm,0)$);
  \end{tikzpicture}

@@

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
    \item<1-> Expose tensor factorization structure through regression.
    \item<2-> Consistent estimator for the mixture of linear regressions with polynomial sample and computational complexity.
    \item<3-> Empirically, method of moment estimates can be a good initialization for EM.
    \item<4-> How can we handle other discriminative models?
      \begin{itemize}
          \item<5-> Dependencies between $h$ and $x$ (mixture of experts).
          \item<6-> Non-linear link functions (logistic regression).
      \end{itemize}
  \end{itemize}
\end{frame}
